---
title: "Quiz"
author: "Siqi Zhu"
date: "December 15, 2015"
output: html_document
---

## Question 1
The American Community Survey distributes downloadable data about United States communities. Download the 2006 microdata survey about housing for the state of Idaho using download.file() from here: 
```
https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv
```
and load the data into R. The code book, describing the variable names is here:
```
https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FPUMSDataDict06.pdf 
```
How many properties are worth $1,000,000 or more?

```{r}
setwd("/Users/zhusiqi/Desktop/coursera/R_jhu/geting_and_cleaning_data/week1")
download.file('https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv','./week1/Q1.csv','curl')
f = read.csv('./Q1.csv')
sum(f$VAL==24 & !is.na(f$VAL))
print(date())
```

## Question 3
Download the Excel spreadsheet on Natural Gas Aquisition Program here: 
```
https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FDATA.gov_NGAP.xlsx 
```
Read rows 18-23 and columns 7-15 into R and assign the result to a variable called:
```
 dat
```
What is the value of:
``` 
sum(dat$Zip*dat$Ext,na.rm=T) 
```
(original data source: http://catalog.data.gov/dataset/natural-gas-acquisition-program)

```{r,results='asis'}
library(xlsx)
download.file('https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FDATA.gov_NGAP.xlsx','./Q3.xlsx','curl')
dat = read.xlsx('./Q3.xlsx',1,rowIndex = 18:23,colIndex = 7:15,header = TRUE)
sum(dat$Zip*dat$Ext,na.rm=T)
print(date())
```

## Question 4
Read the XML data on Baltimore restaurants from here: 
```
https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Frestaurants.xml 
```
How many restaurants have zipcode 21231?

```{r,results='asis'}
library(XML)
doc <-  xmlTreeParse("http://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Frestaurants.xml",useInternal = TRUE)
rootnode <- xmlRoot(doc)
sum(xpathSApply(rootnode,"//zipcode",xmlValue) == 21231)
print(date())
```

## Question 5
The American Community Survey distributes downloadable data about United States communities. Download the 2006 microdata survey about housing for the state of Idaho using download.file() from here: 
```
https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06pid.csv 
```
using the fread() command load the data into an R object
```
 DT
```
Which of the following is the fastest way to calculate the average value of the variable
```
pwgtp15 
```
broken down by sex using the data.table package?

```{r,results='asis'}
library(data.table)
download.file('https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06pid.csv','./Q5.csv','curl')
print(date())
DT = fread('./Q5.csv')
system.time(tapply(DT$pwgtp15,DT$SEX,mean))
system.time(sapply(split(DT$pwgtp15,DT$SEX),mean))
system.time(mean(DT$pwgtp15,by=DT$SEX))
system.time(DT[,mean(pwgtp15),by=SEX])
system.time(mean(DT[DT$SEX==1,]$pwgtp15))
```

